A Linear Solution of the Linearity Challenge

### 2025-12-13 paper/DeadlockFreedom.lagda.tex
revised description of reduction rules
general cleanup
restored deadlock freedom

As we have seen in \Cref{sec:calculus}, the deadlock freedom result
(\Cref{thm:df}) rests on some notions and predicates about processes which we
now formalize in Agda. First of all we need to distinguish \emph{threads} from
\emph{cuts}. As it turns out, in the formalization it is convenient to provide a
more fine-grained classification of threads, distinguishing between links,
input/output actions on linear/shared channels, and sometimes also whether such
actions operate on free or bound channels. We define predicates for each of
these process classes:

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Sum
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Data.List.Base using ([]; _∷_; [_])
open import Relation.Nullary using (¬_; contradiction)
open import Relation.Binary.PropositionalEquality using (refl)

open import Type
open import Context
open import Permutations
open import Process
open import Reduction
open import Congruence
\end{code}

\begin{code}
data Cut            : ∀{Γ} → Process Γ → Set
data Link           : ∀{Γ} → Process Γ → Set
data Input          : ∀{Γ} → Process Γ → Set
data Output         : ∀{Γ} → Process Γ → Set
data Delayed        : ∀{Γ} → Process Γ → Set
data Client         : ∀{Γ} → Process Γ → Set
data Server         : ∀{Γ} → Process Γ → Set
data DelayedServer  : ∀{Γ} → Process Γ → Set
\end{code}

These predicates are relatively uninteresting since they are essentially
isomorphic to the relevant parts of the \AgdaDatatype{Process} datatype. The
only remarkable aspect is that in \AgdaDatatype{Input}, \AgdaDatatype{Output},
\AgdaDatatype{Client} and \AgdaDatatype{Server} the channel being acted upon by
the thread is the \emph{first} in $\Context$, hence it is the \emph{most
recently introduced} channel, whereas in \AgdaDatatype{Delayed} and
\AgdaDatatype{DelayedServer} the channel is older. This allows us to distinguish
those threads that, in the context of a cut, operate on the channel bound by the
cut or on a free channel. As an illustration, consider the
\AgdaInductiveConstructor{wait} case in \AgdaDatatype{Input} and in
\AgdaDatatype{Delayed}. In the former predicate we have

\begin{AgdaAlign}
\begin{code}[hide]
data Cut where
  cut : ∀{Γ Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) -> Cut (cut {A} p P Q)

data Link where
  link : ∀{Γ A} (p : Γ ≃ [ A ] + [ dual A ]) → Link (link p)

data Input where
  fail : ∀{Γ Δ} (p : Γ ≃ [] + Δ) → Input (fail (< p))
\end{code}

\begin{code}
  wait : ∀{Γ Δ P} (p : Γ ≃ [] + Δ) → Input (wait (< p) P)
\end{code}
where the \AgdaInductiveConstructor{<} constructor indicates that the thread
operates on the most recent channel. In the latter predicate we have

\begin{code}[hide]
  case : ∀{Γ Δ A B P Q} (p : Γ ≃ [] + Δ) → Input (case {A} {B} (< p) P Q)
  join : ∀{Γ Δ A B P} (p : Γ ≃ [] + Δ) → Input (join {A} {B} (< p) P)
  all  : ∀{A Γ Δ F} (p : Γ ≃ [] + Δ) -> Input (all {A} (< p) F)

data Output where
  close  : Output close
  left   : ∀{Γ Δ A B P} (p : Γ ≃ [] + Δ) → Output (left {A} {B} (< p) P)
  right  : ∀{Γ Δ A B P} (p : Γ ≃ [] + Δ) → Output (right {A} {B} (< p) P)
  fork   : ∀{Γ Δ Δ₁ Δ₂ A B P Q} (p : Γ ≃ [] + Δ) (q : Δ ≃ Δ₁ + Δ₂) -> Output (fork {A} {B} (< p) q P Q)
  ex     : ∀{A B Γ Δ P} (p : Γ ≃ [] + Δ) -> Output (ex {A} {B} (< p) P)

data Delayed where
  fail     : ∀{C Γ Δ} (p : Γ ∋ ⊤ ⊳ Δ) → Delayed (fail (>_ {C} p))
\end{code}

\begin{code}
  wait     : ∀{C Γ Δ P} (p : Γ ∋ ⊥ ⊳ Δ) → Delayed (wait (>_ {C} p) P)
\end{code}
where the \AgdaInductiveConstructor{>} constructor indicates that the thread
operates on a channel other than the most recent one. Note that here we have to
specify the type $C$ in front of $\Context$ or else Agda is unable to resolve
some metavariables.

\begin{code}[hide]
  case     : ∀{Γ Δ C A B P Q} (p : Γ ∋ A & B ⊳ Δ) → Delayed (case {A} {B} (>_ {C} p) P Q)
  join     : ∀{Γ Δ C A B P} (p : Γ ∋ A ⅋ B ⊳ Δ) → Delayed (join (>_ {C} p) P)
  left     : ∀{Γ Δ C A B P} (p : Γ ∋ A ⊕ B ⊳ Δ) → Delayed (left (>_ {C} p) P)
  right    : ∀{Γ Δ C A B P} (p : Γ ∋ A ⊕ B ⊳ Δ) → Delayed (right (>_ {C} p) P)
  fork-l   : ∀{Γ Δ Δ₁ Δ₂ C A B P Q} (p : Γ ∋ A ⊗ B ⊳ Δ) (q : Δ ≃ Δ₁ + Δ₂) →
             Delayed (fork (>_ {C} p) (< q) P Q)
  fork-r   : ∀{Γ Δ Δ₁ Δ₂ C A B P Q} (p : Γ ∋ A ⊗ B ⊳ Δ) (q : Δ ≃ Δ₁ + Δ₂) →
             Delayed (fork (>_ {C} p) (> q) P Q)
  client   : ∀{Γ Δ A C P} (p : Γ ∋ `? A ⊳ Δ) → Delayed (client (>_ {C} p) P)
  weaken   : ∀{Γ Δ A C P} (p : Γ ∋ `? A ⊳ Δ) → Delayed (weaken (>_ {C} p) P)
  contract : ∀{Γ Δ A C P} (p : Γ ∋ `? A ⊳ Δ) → Delayed (contract (>_ {C} p) P)
  ex       : ∀{A B C Γ Δ P} (p : Γ ∋ `∃ A ⊳ Δ) → Delayed (ex {A} {B} (>_ {C} p) P)
  all      : ∀{A C Γ Δ F} (p : Γ ∋ `∀ A ⊳ Δ) → Delayed (all (>_ {C} p) F)

data Client where
  client   : ∀{Γ Δ A P} (p : Γ ≃ [] + Δ) → Client (client {A} (< p) P)
  weaken   : ∀{Γ Δ A P} (p : Γ ≃ [] + Δ) → Client (weaken {A} (< p) P)
  contract : ∀{Γ Δ A P} (p : Γ ≃ [] + Δ) → Client (contract {A} (< p) P)

data Server where
  server : ∀{Γ Δ A P} (p : Γ ≃ [] + Δ) (un : Un Δ) → Server (server {A} (< p) un P)

data DelayedServer where
  server : ∀{Γ Δ A C P} (p : Γ ∋ `! A ⊳ Δ) (un : Un Δ) →
           DelayedServer (server (> p) (un-∷ {C} un) P)
\end{code}
\end{AgdaAlign}

Now the \AgdaDatatype{Thread} predicate is simply the disjunction of all the
previous ones.

\begin{code}
Thread : ∀{Γ} → Process Γ → Set
Thread P = Link P ⊎ Delayed P ⊎ Output P ⊎ Input P ⊎
           Client P ⊎ Server P ⊎ DelayedServer P
\end{code}

@@

\begin{code}
Observable : ∀{Γ} → Process Γ → Set
Observable P = ∃[ Q ] P ⊒ Q × Thread Q

Reducible : ∀{Γ} → Process Γ → Set
Reducible P = ∃[ Q ] P ↝ Q

Alive : ∀{Γ} → Process Γ → Set
Alive P = Observable P ⊎ Reducible P
\end{code}

\begin{code}[hide]
link-thread : ∀{A Γ} (p : Γ ≃ [ A ] + [ dual A ]) → Thread (link p)
link-thread p = inj₁ (link p)

close-thread : Thread close
close-thread = inj₂ (inj₂ (inj₁ close))

wait-thread : ∀{Γ Δ P} (p : Γ ≃ [ ⊥ ] + Δ) → Thread (wait p P)
wait-thread (< p) = inj₂ (inj₂ (inj₂ (inj₁ (wait p))))
wait-thread (> p) = inj₂ (inj₁ (wait p))

fail-thread : ∀{Γ Δ} (p : Γ ≃ [ ⊤ ] + Δ) → Thread (fail p)
fail-thread (< p) = inj₂ (inj₂ (inj₂ (inj₁ (fail p))))
fail-thread (> p) = inj₂ (inj₁ (fail p))

case-thread : ∀{A B Γ Δ P Q} (p : Γ ≃ [ A & B ] + Δ) → Thread (case p P Q)
case-thread (< p) = inj₂ (inj₂ (inj₂ (inj₁ (case p))))
case-thread (> p) = inj₂ (inj₁ (case p))

join-thread : ∀{A B Γ Δ P} (p : Γ ≃ [ A ⅋ B ] + Δ) → Thread (join p P)
join-thread (< p) = inj₂ (inj₂ (inj₂ (inj₁ (join p))))
join-thread (> p) = inj₂ (inj₁ (join p))

left-thread : ∀{A B Γ Δ P} (p : Γ ∋ A ⊕ B ⊳ Δ) → Thread (left p P)
left-thread (< p) = inj₂ (inj₂ (inj₁ (left p)))
left-thread (> p) = inj₂ (inj₁ (left p))

right-thread : ∀{A B Γ Δ P} (p : Γ ∋ A ⊕ B ⊳ Δ) → Thread (right p P)
right-thread (< p) = inj₂ (inj₂ (inj₁ (right p)))
right-thread (> p) = inj₂ (inj₁ (right p))

fork-thread : ∀{A B Γ Δ Δ₁ Δ₂ P Q} (p : Γ ≃ [ A ⊗ B ] + Δ) (q : Δ ≃ Δ₁ + Δ₂) → Thread (fork p q P Q)
fork-thread (< p) q = inj₂ (inj₂ (inj₁ (fork p q)))
fork-thread (> p) (< q) = inj₂ (inj₁ (fork-l p q))
fork-thread (> p) (> q) = inj₂ (inj₁ (fork-r p q))

client-thread : ∀{A Γ Δ P} (p : Γ ≃ [ `? A ] + Δ) → Thread (client p P)
client-thread (< p) = inj₂ (inj₂ (inj₂ (inj₂ (inj₁ (client p)))))
client-thread (> p) = inj₂ (inj₁ (client p))

weaken-thread : ∀{A Γ Δ P} (p : Γ ≃ [ `? A ] + Δ) → Thread (weaken {A = A} p P)
weaken-thread (< p) = inj₂ (inj₂ (inj₂ (inj₂ (inj₁ (weaken p)))))
weaken-thread (> p) = inj₂ (inj₁ (weaken p))

contract-thread : ∀{A Γ Δ P} (p : Γ ≃ [ `? A ] + Δ) → Thread (contract p P)
contract-thread (< p) = inj₂ (inj₂ (inj₂ (inj₂ (inj₁ (contract p)))))
contract-thread (> p) = inj₂ (inj₁ (contract p))

server-thread : ∀{A Γ Δ P} (p : Γ ≃ [ `! A ] + Δ) (un : Un Δ) → Thread (server p un P)
server-thread (< p) un = inj₂ (inj₂ (inj₂ (inj₂ (inj₂ (inj₁ (server p un))))))
server-thread (> p) (un-∷ un) = inj₂ (inj₂ (inj₂ (inj₂ (inj₂ (inj₂ (server p un))))))

ex-thread : ∀{A B Γ Δ P} (p : Γ ∋ `∃ A ⊳ Δ) -> Thread (ex {A} {B} p P)
ex-thread (< p) = inj₂ (inj₂ (inj₁ (ex p)))
ex-thread (> p) = inj₂ (inj₁ (ex p))

all-thread : ∀{A Γ Δ F} (p : Γ ∋ `∀ A ⊳ Δ) -> Thread (all p F)
all-thread (< p) = inj₂ (inj₂ (inj₂ (inj₁ (all p))))
all-thread (> p) = inj₂ (inj₁ (all p))
\end{code}

@@

\begin{code}
data CanonicalCut {Γ} : Process Γ → Set where
  cc-link           : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                      Link P → CanonicalCut (cut {A} p P Q)
  cc-delayed        : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                      Delayed P → CanonicalCut (cut {A} p P Q)
  cc-delayed-server : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                      DelayedServer P → Server Q → CanonicalCut (cut {A} p P Q)
  cc-redex          : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                      Output P → Input Q → CanonicalCut (cut {A} p P Q)
  cc-connect        : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                      Server P → Client Q → CanonicalCut (cut {A} p P Q)
\end{code}

A \emph{canonical cut} $\Cut[A]\x{P}{Q}$ has one of these forms:
\begin{itemize}
\item $P$ is a link (\AgdaInductiveConstructor{cc-link}), or
\item $P$ operates on a channel other than $x$ and is not a server
(\AgdaInductiveConstructor{cc-delayed}), or
\item $P$ and $Q$ are servers and $P$ operates on a channel other than $x$
(\AgdaInductiveConstructor{cc-delayed-server}), or
\item $P$ performs an output on $x$ and $Q$ performs an input on $x$
(\AgdaInductiveConstructor{cc-redex} and \AgdaInductiveConstructor{cc-connect}).
\end{itemize}

@@

output-client : ∀{A Γ Δ P Q} → ¬ (Output {A ∷ Γ} P × Client {dual A ∷ Δ} Q)
output-client (close , ())

output-server : ∀{A Γ Δ P Q} → ¬ (Output {A ∷ Γ} P × Server {dual A ∷ Δ} Q)
output-server (close , ())

@@

input-client : ∀{A Γ Δ P Q} → ¬ (Input {A ∷ Γ} P × Client {dual A ∷ Δ} Q)
input-client (fail _ , ())

@@

client-client : ∀{A Γ Δ P Q} → ¬ (Client {A ∷ Γ} P × Client {dual A ∷ Δ} Q)
client-client (client _ , ())

client-delayed-server : ∀{A Γ Δ P Q} → ¬ (Client {A ∷ Γ} P × DelayedServer {dual A ∷ Δ} Q)
client-delayed-server (client _ , ())

@@

\begin{code}
canonical-cut : ∀{A Γ Γ₁ Γ₂ P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                Thread P → Thread Q →
                ∃[ R ] CanonicalCut R × cut {A} p P Q ⊒ R
\end{code}

>>>>

As we have seen in \Cref{sec:calculus}, the deadlock freedom result
(\Cref{thm:df}) rests on some notions and predicates about processes which we
now formalize in Agda. First of all we need to define the notion of
\emph{thread}, that is any process other than a cut. As it turns out, it is
convenient to provide an more fine-grained classification of threads,
distinguishing between links, input/output actions, and sometimes also whether
such actions operate on free or bound channels. We define predicates for each of
these process classes:

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Unit using (tt)
open import Data.Sum
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Data.List.Base using ([]; _∷_; [_])
open import Relation.Nullary using (¬_; contradiction)
open import Relation.Unary
open import Relation.Binary.PropositionalEquality using (refl)

open import Type
open import Context
open import Process
open import Reduction
open import Congruence
\end{code}

\begin{code}
data Link           : ∀{Γ} → Proc Γ → Set
data Input          : ∀{Γ} → Proc Γ → Set
data Output         : ∀{Γ} → Proc Γ → Set
data Delayed        : ∀{Γ} → Proc Γ → Set
data Server         : ∀{Γ} → Proc Γ → Set
data DelayedServer  : ∀{Γ} → Proc Γ → Set
\end{code}

These predicates are relatively uninteresting since they are essentially
isomorphic to the relevant parts of the \AgdaDatatype{Proc} datatype. The only
relevant aspect that is worth pointing out here is that in \AgdaDatatype{Input},
\AgdaDatatype{Output} and \AgdaDatatype{Server} the channel being acted upon by
the thread is the \emph{first} in $\Context$, hence it is the \emph{most
recently introduced} channel, whereas in \AgdaDatatype{Delayed} and
\AgdaDatatype{DelayedServer} the channel is older. This allows us to distinguish
those threads that, in the context of a cut, operate on the channel bound by the
cut or on a free channel. As an illustration, consider the
\AgdaInductiveConstructor{wait} case in \AgdaDatatype{Input} and in
\AgdaDatatype{Delayed}. In the former predicate we have

\begin{AgdaAlign}
\begin{code}[hide]
data Link where
  link : ∀{Γ A} (p : Γ ≃ [ A ] + [ dual A ]) → Link (link (ch ⟨ p ⟩ ch))

data Input where
  fail : ∀{Γ Δ} (p : Γ ≃ [] + Δ) → Input (fail (ch ⟨ < p ⟩ tt))
\end{code}

\begin{code}
  wait : ∀{Γ Δ P} (p : Γ ≃ [] + Δ) → Input (wait (ch ⟨ < p ⟩ P))
\end{code}
where the use of the \AgdaInductiveConstructor{<} constructor indicates that the
thread operates on the most recent channel. In the latter predicate we have

\begin{code}[hide]
  case : ∀{Γ Δ A B P Q} (p : Γ ≃ [] + Δ) → Input (case {A} {B} (ch ⟨ < p ⟩ (P , Q)))
  join : ∀{Γ Δ A B P} (p : Γ ≃ [] + Δ) → Input (join {A} {B} (ch ⟨ < p ⟩ P))
  all  : ∀{A Γ Δ F} (p : Γ ≃ [] + Δ) → Input (all {A} (ch ⟨ < p ⟩ F))

data Output where
  close    : Output (close ch)
  select-l : ∀{Γ Δ A B P} (p : Γ ≃ [] + Δ) → Output (select {A} {B} (ch ⟨ < p ⟩ inj₁ P))
  select-r : ∀{Γ Δ A B P} (p : Γ ≃ [] + Δ) → Output (select {A} {B} (ch ⟨ < p ⟩ inj₂ P))
  fork     : ∀{Γ Δ Δ₁ Δ₂ A B P Q} (p : Γ ≃ [] + Δ) (q : Δ ≃ Δ₁ + Δ₂) → Output (fork {A} {B} (ch ⟨ < p ⟩ (P ⟨ q ⟩ Q)))
  ex       : ∀{A B Γ Δ P} (p : Γ ≃ [] + Δ) → Output (ex {A} {B} (ch ⟨ < p ⟩ P))
  client   : ∀{Γ Δ A P}   (p : Γ ≃ [] + Δ) → Output (client {A} (ch ⟨ < p ⟩ P))
  weaken   : ∀{Γ Δ A P}   (p : Γ ≃ [] + Δ) → Output (weaken {A} (ch ⟨ < p ⟩ P))
  contract : ∀{Γ Δ A P}   (p : Γ ≃ [] + Δ) → Output (contract {A} (ch ⟨ < p ⟩ P))

data Delayed where
  fail     : ∀{C Γ Δ} (p : Γ ≃ [ ⊤ ] + Δ) → Delayed (fail (ch ⟨ >_ {C} p ⟩ tt))
\end{code}

\begin{code}
  wait     : ∀{C Γ Δ P} (p : Γ ≃ [ ⊥ ] + Δ) → Delayed (wait (ch ⟨ >_ {C} p ⟩ P))
\end{code}
where the use of the \AgdaInductiveConstructor{>} constructor indicates that the
thread operates on a channel other than the most recent one. Note that here we
have to specify the type $C$ in front of $\Context$ or else Agda is unable to
resolve some metavariables.

\begin{code}[hide]
  case     : ∀{Γ Δ C A B P} (p : Γ ≃ [ A & B ] + Δ) → Delayed (case {A} {B} (ch ⟨ >_ {C} p ⟩ P))
  select-l : ∀{Γ Δ C A B P} (p : Γ ≃ [ A ⊕ B ] + Δ) → Delayed (select (ch ⟨ >_ {C} p ⟩ inj₁ P))
  select-r : ∀{Γ Δ C A B P} (p : Γ ≃ [ A ⊕ B ] + Δ) → Delayed (select (ch ⟨ >_ {C} p ⟩ inj₂ P))
  join     : ∀{Γ Δ C A B P} (p : Γ ≃ [ A ⅋ B ] + Δ) → Delayed (join (ch ⟨ >_ {C} p ⟩ P))
  fork-l   : ∀{Γ Δ Δ₁ Δ₂ C A B P Q} (p : Γ ≃ [ A ⊗ B ] + Δ) (q : Δ ≃ Δ₁ + Δ₂) →
             Delayed (fork (ch ⟨ >_ {C} p ⟩ (P ⟨ < q ⟩ Q)))
  fork-r   : ∀{Γ Δ Δ₁ Δ₂ C A B P Q} (p : Γ ≃ [ A ⊗ B ] + Δ) (q : Δ ≃ Δ₁ + Δ₂) →
             Delayed (fork (ch ⟨ >_ {C} p ⟩ (P ⟨ > q ⟩ Q)))
  all      : ∀{A C Γ Δ F} (p : Γ ≃ [ `∀ A ] + Δ) → Delayed (all (ch ⟨ >_ {C} p ⟩ F))
  ex       : ∀{A B C Γ Δ P} (p : Γ ≃ [ `∃ A ] + Δ) → Delayed (ex {A} {B} (ch ⟨ >_ {C} p ⟩ P))
  client   : ∀{Γ Δ A C P} (p : Γ ≃ [ `? A ] + Δ) → Delayed (client (ch ⟨ >_ {C} p ⟩ P))
  weaken   : ∀{Γ Δ A C P} (p : Γ ≃ [ `? A ] + Δ) → Delayed (weaken (ch ⟨ >_ {C} p ⟩ P))
  contract : ∀{Γ Δ A C P} (p : Γ ≃ [ `? A ] + Δ) → Delayed (contract (ch ⟨ >_ {C} p ⟩ P))

data Server where
  server : ∀{Γ Δ A P} (p : Γ ≃ [] + Δ) (un : Un Δ) → Server (server {A} (ch ⟨ < p ⟩ (un , P)))

data DelayedServer where
  server : ∀{Γ Δ A C P} (p : Γ ≃ [ `! A ] + Δ) (un : Un Δ) →
           DelayedServer (server {A} (ch ⟨ >_ p ⟩ (un-∷ {C} un , P)))
\end{code}
\end{AgdaAlign}

Now the \AgdaDatatype{Thread} predicate is simply the disjoint union of all the
previous ones.

\begin{code}
data Thread {Γ} (P : Proc Γ) : Set where
  link     : Link P → Thread P
  delayed  : Delayed P → Thread P
  output   : Output P → Thread P
  input    : Input P → Thread P
  server   : Server P → Thread P
  dserver  : DelayedServer P → Thread P
\end{code}

@@

\begin{code}
Observable : ∀{Γ} → Proc Γ → Set
Observable P = ∃[ Q ] P ⊒ Q × Thread Q

Reducible : ∀{Γ} → Proc Γ → Set
Reducible P = ∃[ Q ] P ↝ Q

Alive : ∀{Γ} → Proc Γ → Set
Alive P = Observable P ⊎ Reducible P
\end{code}

\begin{code}[hide]
fail→thread : ∀{Γ Δ} (p : Γ ≃ [ ⊤ ] + Δ) → Thread (fail (ch ⟨ p ⟩ tt))
fail→thread (< p) = input (fail p)
fail→thread (> p) = delayed (fail p)

wait→thread : ∀{Γ Δ P} (p : Γ ≃ [ ⊥ ] + Δ) → Thread (wait (ch ⟨ p ⟩ P))
wait→thread (< p) = input (wait p)
wait→thread (> p) = delayed (wait p)

case→thread : ∀{A B Γ Δ P} (p : Γ ≃ [ A & B ] + Δ) → Thread (case (ch ⟨ p ⟩ P))
case→thread (< p) = input (case p)
case→thread (> p) = delayed (case p)

left→thread : ∀{A B Γ Δ P} (p : Γ ≃ [ A ⊕ B ] + Δ) → Thread (select (ch ⟨ p ⟩ inj₁ P))
left→thread (< p) = output (select-l p)
left→thread (> p) = delayed (select-l p)

right→thread : ∀{A B Γ Δ P} (p : Γ ≃ [ A ⊕ B ] + Δ) → Thread (select (ch ⟨ p ⟩ inj₂ P))
right→thread (< p) = output (select-r p)
right→thread (> p) = delayed (select-r p)

join→thread : ∀{A B Γ Δ P} (p : Γ ≃ [ A ⅋ B ] + Δ) → Thread (join (ch ⟨ p ⟩ P))
join→thread (< p) = input (join p)
join→thread (> p) = delayed (join p)

fork→thread : ∀{A B Γ Δ Δ₁ Δ₂ P Q} (p : Γ ≃ [ A ⊗ B ] + Δ) (q : Δ ≃ Δ₁ + Δ₂) → Thread (fork (ch ⟨ p ⟩ (P ⟨ q ⟩ Q)))
fork→thread (< p) q = output (fork p q)
fork→thread (> p) (< q) = delayed (fork-l p q)
fork→thread (> p) (> q) = delayed (fork-r p q)

all→thread : ∀{A Γ Δ F} (p : Γ ≃ [ `∀ A ] + Δ) → Thread (all (ch ⟨ p ⟩ F))
all→thread (< p) = input (all p)
all→thread (> p) = delayed (all p)

ex→thread : ∀{A B Γ Δ P} (p : Γ ≃ [ `∃ A ] + Δ) → Thread (ex {A} {B} (ch ⟨ p ⟩ P))
ex→thread (< p) = output (ex p)
ex→thread (> p) = delayed (ex p)

server→thread : ∀{A Γ Δ P} (p : Γ ≃ [ `! A ] + Δ) (un : Un Δ) → Thread (server (ch ⟨ p ⟩ (un , P)))
server→thread (< p) un = server (server p un)
server→thread (> p) (un-∷ un) = dserver (server p un)

client→thread : ∀{A Γ Δ P} (p : Γ ≃ [ `? A ] + Δ) → Thread (client (ch ⟨ p ⟩ P))
client→thread (< p) = output (client p)
client→thread (> p) = delayed (client p)

weaken→thread : ∀{A Γ Δ P} (p : Γ ≃ [ `? A ] + Δ) → Thread (weaken {A = A} (ch ⟨ p ⟩ P))
weaken→thread (< p) = output (weaken p)
weaken→thread (> p) = delayed (weaken p)

contract→thread : ∀{A Γ Δ P} (p : Γ ≃ [ `? A ] + Δ) → Thread (contract (ch ⟨ p ⟩ P))
contract→thread (< p) = output (contract p)
contract→thread (> p) = delayed (contract p)
\end{code}

@@

\begin{code}
data CanonicalCut {Γ} : Proc Γ → Set where
  cc-link     : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                Link P → CanonicalCut (cut {A} (P ⟨ p ⟩ Q))
  cc-redex    : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                Output P → (Input ∪ Server) Q →
                CanonicalCut (cut {A} (P ⟨ p ⟩ Q))
  cc-delayed  : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                Delayed P → CanonicalCut (cut {A} (P ⟨ p ⟩ Q))
  cc-servers  : ∀{Γ₁ Γ₂ A P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                DelayedServer P → Server Q →
                CanonicalCut (cut {A} (P ⟨ p ⟩ Q))
\end{code}

A \emph{canonical cut} $\Cut[A]\x{P}{Q}$ has one of these forms:
\begin{itemize}
\item $P$ is a link (\AgdaInductiveConstructor{cc-link}), or
\item $P$ performs an output on $x$ and $Q$ performs an input on $x$
  (\AgdaInductiveConstructor{cc-redex}), or
\item $P$ operates on a channel other than $x$ and is not a server
(\AgdaInductiveConstructor{cc-delayed}), or
\item $P$ and $Q$ are servers and $P$ operates on a channel other than $x$
(\AgdaInductiveConstructor{cc-servers}).
\end{itemize}

Note that we have to distinguish servers from the other input operations because
the structural pre-congruence rule \SServer can only be applied when the two
sub-processes of a cut are both servers.

@@

@@

@@

@@

\begin{code}
canonical-cut : ∀{A Γ Γ₁ Γ₂ P Q} (p : Γ ≃ Γ₁ + Γ₂) →
                Thread P → Thread Q →
                ∃[ R ] CanonicalCut R × cut {A} (P ⟨ p ⟩ Q) ⊒ R
\end{code}

\begin{code}[hide]
canonical-cut pc (link x) Qt = _ , cc-link pc x , s-refl
canonical-cut pc Pt (link y) = _ , cc-link (+-comm pc) y , s-comm pc
canonical-cut pc (delayed x) Qt = _ , cc-delayed pc x , s-refl
canonical-cut pc Pt (delayed y) = _ , cc-delayed (+-comm pc) y , s-comm pc
canonical-cut pc (output x) (output y) = contradiction (x , y) output-output
canonical-cut pc (output x) (input y) = _ , cc-redex pc x (inj₁ y) , s-refl
canonical-cut pc (output x) (server y) = _ , cc-redex pc x (inj₂ y) , s-refl
canonical-cut pc (output x) (dserver y) = contradiction (x , y) output-delayed-server
canonical-cut pc (input x) (output y) = _ , cc-redex (+-comm pc) y (inj₁ x) , s-comm pc
canonical-cut pc (input x) (input y) = contradiction (x , y) input-input
canonical-cut pc (input x) (server y) = contradiction (x , y) input-server
canonical-cut pc (input x) (dserver y) = contradiction (x , y) input-delayed-server
canonical-cut pc (server x) (output y) = _ , cc-redex (+-comm pc) y (inj₂ x) , s-comm pc
canonical-cut pc (server x) (input y) = contradiction (y , x) input-server
canonical-cut pc (server x) (server y) = contradiction (x , y) server-server
canonical-cut pc (server x) (dserver y) = _ , cc-servers (+-comm pc) y x , s-comm pc
canonical-cut pc (dserver x) (output y) = contradiction (y , x) output-delayed-server
canonical-cut pc (dserver x) (input y) = contradiction (y , x) input-delayed-server
canonical-cut pc (dserver x) (server y) = _ , cc-servers pc x y , s-refl
canonical-cut pc (dserver x) (dserver y) = contradiction (x , y) delayed-server-delayed-served

⊒Alive : ∀{Γ} {P Q : Proc Γ} → P ⊒ Q → Alive Q → Alive P
⊒Alive pcong (inj₁ (_ , x , th)) = inj₁ (_ , s-tran pcong x , th)
⊒Alive pcong (inj₂ (_ , red)) = inj₂ (_ , r-cong pcong red)
\end{code}
